Downstream analysis: A pipeline to analyse time-course gene expression data

### 2019-05-08 scripts/01_quality_control_normalization.Rmd
Separating references from the rest of the QC
More citations, more text

```{r message=FALSE}
targets = limma::readTargets()
RG = limma::read.maimages(
  targets, source="agilent",
  path=data_dir,
  other.columns="gIsWellAboveBG",
  green.only=TRUE)

@@

```{r reading_targets, results="hide", fig.width=8, fig.height=4, message=FALSE}
# Perform background correction using norm-exponential strategy and normalize
# across samples using quantile normalization.
RG = limma::backgroundCorrect(RG, method="normexp")
RG.q = limma::normalizeBetweenArrays(RG, method="quantile")
```

@@

```{r plotting_densities, results="hide", fig.width=8, fig.height=4}
# Plot the densities of the probes before and after normalization.
par(mfrow=c(1, 2))
limma::plotDensities(RG, col="black", legend=FALSE)
limma::plotDensities(RG.q, col="black", legend=FALSE)
```

@@

```{r}
# Now load in the metadata
data(shoemaker2015)
meta = shoemaker2015$meta
```

>>>>

```{r message=FALSE, results="hide", warning=FALSE}
targets = limma::readTargets()
RG = limma::read.maimages(
  targets, source="agilent",
  path=data_dir,
  other.columns="gIsWellAboveBG",
  green.only=TRUE)

@@

```{r reading_targets, results="hide", fig.width=8, fig.height=4, message=FALSE, warning=FALSE}
# Perform background correction using norm-exponential strategy and normalize
# across samples using quantile normalization.
RG = limma::backgroundCorrect(RG, method="normexp")
RG.q = limma::normalizeBetweenArrays(RG, method="quantile")
```

@@

```{r plotting_densities, results="hide", fig.width=8, fig.height=4}
# Plot the densities of the probes before and after normalization.
par(mfrow=c(1, 2))
limma::plotDensities(RG, col="black", legend=FALSE,
		     main="Before normalization")
limma::plotDensities(RG.q, col="black", legend=FALSE,
		     main="After normalization")
```

@@

```{r}
# Now load in the metadata
data(shoemaker2015)
meta = shoemaker2015$meta
data = shoemaker2015$data
```

### 2019-05-08 scripts/03_clustering.Rmd
More citations, more text

```{r echo=FALSE, fig.height=6, fig.width=6}
par(mfrow=c(2, 1))
barplot(table(kmeans_clusters$clusters), col="black", main="splines k-means
labeling")
barplot(table(labels), col="black", main="score and label")
```

@@

```{r elbow_method, fig.width=6, fig.height=4}

# function to compute total within-cluster sum of square

plot(all_possible_n_clusters, wss_values,
     type="b", pch=19, frame=FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")

@@

>>>>

```{r echo=FALSE, fig.height=6, fig.width=6, warning=FALSE}
par(mfrow=c(2, 1))
barplot(table(kmeans_clusters$clusters), col="black", main="splines k-means
labeling")
barplot(table(labels), col="black", main="score and label")
```

@@

The Elbow method to choose the number of clusters relies on visualization aid
to choose the number of cluster. The method relies on plotting the within
cluster sum of squares as a function of the number of clusters. At some point,
the WCSS will stop dropping, giving an angle in the graph.  The number of
cluster is chosen at this "Elbow point."

```{r elbow_method, fig.width=6, fig.height=4}
plot(all_possible_n_clusters, wss_values,
     type="b", pch=19, frame=FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")

@@

The silhouette value is a measure of how similar a data point is to its own
cluster (cohesion) compared to other clusters (separation).

### 2019-05-08 scripts/04_downstream_analysis.Rmd
More and better text. also better k for clustering
More and better texts
More citations, more text

---
title: "Downstream analysis: A pipeline to analyse time-course gene expression data"
output: rmarkdown::html_vignette
---

@@

Let us first tackle the case of pathway enrichment analysis. We will leverage
the packages `biomaRt` and `KEGGprofile` for this step. `KEGGprofile` is a
package that easily allows to perform pathway enrichment analysis on a set of
genes labeled with the ensembl annotation. We thus need to convert the gene
names into the appropriate format. This is where `biomaRt` comes in handy: it
enables easy conversion from one gene annotation to another. Here, we will use
it to convert the gene names from the Refseq annotation to the ensembl one.

@@

```{r enrichment_analysis, results="hide", message=FALSE}
cluster = 20
genes = gene_names[labels == cluster]

@@

colnames(pathways) = c("Pathway", "Percentage", "Adj. p-value")
pathways["Percentage"] = pathways["Percentage"] * 100
kable(pathways, "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```

@@

# Create gene to GO id mapping
gene_id_go_mapping = create_go_term_mapping(genes)
```

@@

kable(go_terms_enriched, "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```

>>>>

---
title: "Downstream analysis: A pipeline to analyse time-course gene expression data"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
---

@@

Let us first tackle the case of pathway enrichment analysis. We will leverage
the packages `biomaRt` [@durink:biomart] and `KEGGprofile` [@zhao:keggprofile]
for this step. `KEGGprofile` is a
package that easily allows to perform pathway enrichment analysis on a set of
genes labeled with the ensembl annotation. We thus need to convert the gene
names into the appropriate format. This is where `biomaRt` comes in handy: it
enables easy conversion from one gene annotation to another. Here, we will use
it to convert the gene names from the Refseq annotation to the ensembl one.

@@

```{r enrichment_analysis, results="hide", message=FALSE}
cluster = 8
genes = gene_names[labels == cluster]

@@

colnames(pathways) = c("Pathway", "Percentage", "Adj. p-value")
pathways["Percentage"] = pathways["Percentage"] * 100
kable(pathways[1:10, ], "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```

@@

# Create gene to GO id mapping
gene_id_go_mapping = moanin::create_go_term_mapping(genes)
```

@@

kable(go_terms_enriched[1:15, ], "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```

# References
